Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-04-03 thesis/chapters/ch_chapter3.tex
Updated references and new refences for ch4

\subsection{Implications for Exotic Enrichment Sources}
\label{ch3:sec:exotic enrichment}
%   AE: Benoit suggested removing arguments relating to NS-NS merger energetics as I had misred / miscaclulated the energy release values. He said greater emphasise on potential differences of mixing due to rarity, rather than energy.
We have shown that the dynamical evolution of metals depends upon their nucleosynthetic source, focusing here on AGB synthesized elements and SN synthesized elements. However, these differences should also apply to exotic enrichment sources, such as hypernovae, neutron-star neutron-star mergers, and neutron-star black-hole mergers. These sources can have energies that differ significantly from typical SNe, reaching $> 10^{52}$~erg for hypernovae \citep{Nomoto2004}, for example. In addition, 
%mm the rarity of these events, for example,  there are approximately $10^{-5}$ per M$_{\odot}$ of star formation NSM in the Milky Way, could
     these events are rare.  For example, neutron star mergers occur 
     at a rate of approximately $10^{-5}$ per M$_{\odot}^{-1}$ of star
     formation in the Milky Way\citep{Kim2015}.  This rarity could 
significantly influence the typical ISM environments into which they eject their metals. We would therefore expect different mixing behaviors for metals synthesized through these channels as compared to SNe. Based on our results here, for example, we would expect elements from hypernovae to be more well mixed in the ISM of all galaxies, but more readily ejected in dwarf galaxies, as compared to elements from SNe. 
% Benoit suggesed removing this given energy argument removal above
%    Elements from NSMs, a significant source of r-process enrichment, would be less well-mixed in the ISM of all galaxies and more easily retained than SNe elements in dwarf galaxies. 
Depending on their injection energy, neutron star mergers could exhibit different mixing behaviors in the ISM. These differences could provide important signatures for distinguishing individual, exotic enrichment sources from observed stellar abundance patterns in our own Milky Way and in nearby dwarf galaxies. In particular, low mass dwarf galaxies with unusual (compared to Milky Way, solar abundances) r-process enrichment \citep[e.g.][]{Ji2016,Ji2018b,Duggan2018}, are valuable for constraining the source of these elements, their frequency, and typical yields. The differential metal evolution presented in our work both opens up an additional avenue by which elements from distinct nucleosynthetic sources may be distinguished in observations and challenges current assumptions used in interpreting these observations.

>>>>

\subsection{Implications for Exotic Enrichment Sources}
\label{ch3:sec:exotic enrichment}
%   AE: Benoit suggested removing arguments relating to NS-NS merger energetics as I had misred / miscaclulated the energy release values. He said greater emphasise on potential differences of mixing due to rarity, rather than energy.
We have shown that the dynamical evolution of metals depends upon their nucleosynthetic source, focusing here on AGB synthesized elements and SN synthesized elements. However, these differences should also apply to exotic enrichment sources, such as hypernovae, neutron-star neutron-star mergers, and neutron-star black-hole mergers. These sources can have energies that differ significantly from typical SNe, reaching $> 10^{52}$~erg for hypernovae \citep{Nomoto2004}, for example. In addition,
%mm the rarity of these events, for example,  there are approximately $10^{-5}$ per M$_{\odot}$ of star formation NSM in the Milky Way, could
     these events are rare.  For example, neutron star mergers occur
     at a rate of approximately $10^{-5}$ per M$_{\odot}^{-1}$ of star
     formation in the Milky Way\citep{Kim2015}.  This rarity could
significantly influence the typical ISM environments into which they eject their metals. We would therefore expect different mixing behaviors for metals synthesized through these channels as compared to SNe. Based on our results here, for example, we would expect elements from hypernovae to be more well mixed in the ISM of all galaxies, but more readily ejected in dwarf galaxies, as compared to elements from SNe.
% Benoit suggesed removing this given energy argument removal above
%    Elements from NSMs, a significant source of r-process enrichment, would be less well-mixed in the ISM of all galaxies and more easily retained than SNe elements in dwarf galaxies.
Depending on their injection energy, neutron star mergers could exhibit different mixing behaviors in the ISM. These differences could provide important signatures for distinguishing individual, exotic enrichment sources from observed stellar abundance patterns in our own Milky Way and in nearby dwarf galaxies. In particular, low mass dwarf galaxies with unusual (compared to Milky Way, solar abundances) r-process enrichment \citep[e.g.][]{Ji2016,Duggan2018,Ji2019}, are valuable for constraining the source of these elements, their frequency, and typical yields. The differential metal evolution presented in our work both opens up an additional avenue by which elements from distinct nucleosynthetic sources may be distinguished in observations and challenges current assumptions used in interpreting these observations.

### 2019-04-03 thesis/chapters/ch_introduction.tex
Updated references and new refences for ch4

The observational landscape today is primed for developing a more detailed understanding these processes. Recent observational campaigns such as SEGUE \citep{Yanny2009}, RAVE \citep{Kunder2017}, the Gaia-ESO survey \citep{Gaia}, APOGEE and APOGEE2 \citep{APOGEE2010,APOGEE}, GALAH \citep{GALAH,Buder2018}, as well as upcoming observations, such as the Local Volume Mapper as part of SDSS-V, have generated tremendous amounts of information on detailed stellar abundances and stellar kinematics in our Milky Way and nearby Local Group dwarf galaxies. One of the most powerful proposed uses of this enormous trove of data is chemical tagging \citep{Freeman2002}, whereby stellar populations are analyzed in chemical space and 6D phase space to identify co-eval and co-natal groups of stars. This process of galactic archeology aims to break down and identify each distinct stellar component of our Galaxy, explaining the process of their formation and evolution. Substantial work has been made recently to determine the efficacy of this approach \citep[e.g.][]{Ting2012,Hogg2016,Jofre107,Price-Jones2018,Armillotta2018}, yet the physical processes that give rise to stellar abundances as we observe them today are still uncertain.

@@

Radiative processes are of fundamental importance in properly modeling the diversity of density, temperature, and ionization states in a multi-phase ISM. It is via radiative cooling that gas, heated by gravitational accretion onto dark matter halos, can cool, condense, become self-gravitating, and (eventually) collapse to form stars. Conversely heating, via radiative absorption or scatterings, from stellar sources within galaxies and the extragalactic sources that comprises the cosmic UV background (UVB, e.g. \cite{HM2001,HM2012,FG2011}). Tied to both of these processes is the actual chemistry that occurs in the ISM, from the primordial chemical reactions between H, He, and free electrons to substantially more complicated reactions that occur once significant amounts of metals are present in the ISM (particularly, C, O, and Si). These metals by themselves act as additional radiative coolants, which strongly dictate the shape of the cooling curve that helps establish a multi-phase medium in the first place \citep[e.g.][]{McKeeOstriker1977}. The molecules produced in chemical reactions with these metals produce additional coolants in the ISM \citep{HollenbackMcKee1979}, including dust, which has its own rich array of associated physical processes \citep{Omukai2000,Omukai2005,Draine2011}.

A full, detailed model of radiative cooling, heating, non-equillibrium chemistry, and dust is generally far too computationally expensive to include in galaxy-scale simulations. For this reason, many large-scale cosmological simulations assume that all gas is optically thin and that all molecules and ionization states exist in equilibrium. This allows for the quick use of tabulated look-up tables for radiative cooling and heating, often produced using \code{cloudy} \citep{Cloudy2013}, that depend on gas density, temperature, metallicity, and redshift (which sets the UVB heating rate). More complex simulations often track some number (usually $< 10$) of species involved in primordial chemistry in order to better account for non-equilibrium effects and \Hmolecular production, though with noted increases in memory requirements and computational cost. Some of the most advance models to-date can account for well over 10$^2$ individual chemical reactions involving $>$ 20 elements and molecules \citep[e.g.][]{Richings2015,Richings2018}; but these are typically prohibitively expensive to run on large scales. In this \dissertation we adopt the chemistry and radiative cooling and heating physics models available in the newly developed, open-source library \code{GRACKLE} \citep{GrackleMethod}. The exact physics included in our simulations are discussed in greater detail in Chapter~\ref{ch:chapter1}, but we note that improvements to \code{GRACKLE} have been made as the direct result of the research conducted during this \dissertation.
%mention that I made improvements to this?

>>>>

The observational landscape today is primed for developing a more detailed understanding these processes. Recent observational campaigns such as SEGUE \citep{Yanny2009}, RAVE \citep{Kunder2017}, the Gaia-ESO survey \citep{Gaia}, APOGEE and APOGEE2 \citep{APOGEE2010,APOGEE}, GALAH \citep{GALAH,Buder2018}, as well as upcoming observations, such as the Local Volume Mapper as part of SDSS-V, have generated tremendous amounts of information on detailed stellar abundances and stellar kinematics in our Milky Way and nearby Local Group dwarf galaxies. One of the most powerful proposed uses of this enormous trove of data is chemical tagging \citep{Freeman2002}, whereby stellar populations are analyzed in chemical space and 6D phase space to identify co-eval and co-natal groups of stars. This process of galactic archeology aims to break down and identify each distinct stellar component of our Galaxy, explaining the process of their formation and evolution. Substantial work has been made recently to determine the efficacy of this approach \citep[e.g.][]{Ting2012,Hogg2016,Jofre2017,Price-Jones2018,Armillotta2018}, yet the physical processes that give rise to stellar abundances as we observe them today are still uncertain.

@@

Radiative processes are of fundamental importance in properly modeling the diversity of density, temperature, and ionization states in a multi-phase ISM. It is via radiative cooling that gas, heated by gravitational accretion onto dark matter halos, can cool, condense, become self-gravitating, and (eventually) collapse to form stars. Conversely heating, via radiative absorption or scatterings, from stellar sources within galaxies and the extragalactic sources that comprises the cosmic UV background (UVB, e.g. \cite{HM2001,HM2012,FG2011}). Tied to both of these processes is the actual chemistry that occurs in the ISM, from the primordial chemical reactions between H, He, and free electrons to substantially more complicated reactions that occur once significant amounts of metals are present in the ISM (particularly, C, O, and Si). These metals by themselves act as additional radiative coolants, which strongly dictate the shape of the cooling curve that helps establish a multi-phase medium in the first place \citep[e.g.][]{McKeeOstriker1977}. The molecules produced in chemical reactions with these metals produce additional coolants in the ISM \citep{HollenbachMcKee1979}, including dust, which has its own rich array of associated physical processes \citep{Omukai2000,Omukai2005,Draine2011}.

A full, detailed model of radiative cooling, heating, non-equillibrium chemistry, and dust is generally far too computationally expensive to include in galaxy-scale simulations. For this reason, many large-scale cosmological simulations assume that all gas is optically thin and that all molecules and ionization states exist in equilibrium. This allows for the quick use of tabulated look-up tables for radiative cooling and heating, often produced using \code{cloudy} \citep{Cloudy2013}, that depend on gas density, temperature, metallicity, and redshift (which sets the UVB heating rate). More complex simulations often track some number (usually $< 10$) of species involved in primordial chemistry in order to better account for non-equilibrium effects and \Hmolecular production, though with noted increases in memory requirements and computational cost. Some of the most advance models to-date can account for well over 10$^2$ individual chemical reactions involving $>$ 20 elements and molecules \citep[e.g.][]{Richings2016,Richings2018}; but these are typically prohibitively expensive to run on large scales. In this \dissertation we adopt the chemistry and radiative cooling and heating physics models available in the newly developed, open-source library \code{GRACKLE} \citep{GrackleMethod}. The exact physics included in our simulations are discussed in greater detail in Chapter~\ref{ch:chapter1}, but we note that improvements to \code{GRACKLE} have been made as the direct result of the research conducted during this \dissertation.
%mention that I made improvements to this?
